---
title: "To my students learning R: these are some of the mistakes I made"
author: Tom Palmer
date: '2022-12-09'
slug: learning-r-mistakes
categories:
  - Blog
tags:
  - R
subtitle: ''
summary: ''
authors: []
lastmod: '2022-12-03T10:37:33Z'
featured: no
draft: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r}
knitr::opts_chunk$set(collapse = TRUE)
```

## Introduction

Are you learning R? Let's cover a few R features/pitfalls/gotchas that you might come across.

## Vectors are recycled

Recycling rules in base R are slightly inconsistent.^[See [here](https://vctrs.r-lib.org/articles/type-size.html#appendix-recycling-in-base-r)]

## Assignment versus setting arguments

We can perform left assignment with either `<-` or `=`, however we set argument values with only `=`.

But additionally we can perform assignment within another function call.




## Vectors are recycled

## Objects are overwritten without warning

```{r}
myvariable <- 10
myvariable <- "Hello, world!"
myvariable
```

## Computers store numbers in different ways

```{r}
class(1)
class(1L)
```

## Different packages with functions with the same name

Namespace conflict


My advice, load the **tidyverse** last in your list of `library()` calls.
